\chapter{Introduction}
\label{cha:introduction}

Propelled by the increasing usage of web enabled, context and location
aware mobile devices, context aware services specially location based
services (LBS) are becoming remarkably popular. The recent
developments seen in technologies like GPS, GSM, Wifi and Bluetooth
also worked as a catalyst in this process. Using these technologies it
is possible to obtain information about the location of users. This
type of information can be used both by service providers and
users. Service providers can process this information to obtain a
better understanding of the users behavior (e.g. by analyzing their
movement) and improve their current services, as well as plans for
future services and infrastructures. Users, on the other hand, can use
location data to get personalized information according to their
location. However, the continuous monitoring, processing and storage
of location data can create privacy related problems. Location
information can expose sensitive information about users as it
constitutes a \emph{quasi-identifier}\footnote{Data that by itself
  does not uniquely identify a user but that in association with other
  information might.} according to Bettini et
al.~\cite{bettini2005protecting}.

Examples of attacks on privacy using recorded location data are not
hard to find. Using a week-long database, with GPS traces of Detroit
drivers whose names were replaced with pseudonyms, Hoh et
al.~\cite{hoh2006enhancing} were able to determine the location of
85\% of the drivers' homes. Hightower et al.'s BeaconPrint
algorithm~\cite{hightower2005learning} is able to accurately recognize
the location of a user using only Wifi and GSM
information. Researchers have also shown that, looking at the historic
of someone's location data, it is possible to predict their future
whereabouts~\cite{krumm2006predestination,froehlich2008route}.

While location based information is very useful, the fact that it can
be used to attack the privacy of the users might be a decisive factor
in the popularity and sustainability of services supported by this
type of information.  The solution to this problem consists in the use
of privacy preserving approaches capable of ensuring both the quality
of the service and the privacy of its users.

Stochastic summarizing techniques are space efficient, probabilistic
mechanisms, capable of producing summaries of the data they receive as
input.  The size of the summaries depends only on the number of input
elements, not on the size of the elements. Being non-deterministic,
these techniques do not allow the original item to be recreated from
its resulting summary. However they do offer a trade-of between space
and accuracy where one is obtained at the expense of the other.

% i.e., it is always $x * N$ bits, where $x$ is a
% predetermined value and $N$ is the number of elements

This dissertation aims to find if stochastic summarizing techniques
are suitable as privacy preserving techniques regarding location
data in large scale scenarios. As a proof of concept, we present two
Bluetooth based scenarios, where we make use of these techniques
to ensure the privacy of the users while trying to comply with the
scenarios' requirements.

Bluetooth technology is a relatively easy and inexpensive way to track
large groups of people. Unlike technologies like GPS, Bluetooth also
works well in both indoors and outdoors scenarios. This allows the
tracking of individuals both inside and outside buildings allowing for
more robust tracking solutions.  Also, Bluetooth has a limited
detection range (typically between 10 to 100 meters depending on the
number of obstacles) which makes the differentiation between nearby places
possible, thus enabling accurate tracking results. Furthermore,
Bluetooth tracking is a passive tracking process, i.e, it does not
require users to change their behavior. On the other hand, GPS
positions are computed in the mobile device and have to be handed to
the tracking entity.  All these properties make Bluetooth an arguably
good choice in which to base our tracking scenarios.

As Bluetooth becomes more and more pervasive, there is a growing
potential to leverage on the possibilities offered by Bluetooth
scanning as a flexible infrastructure for situated interaction and a
general purpose platform for massive sensing and actuation in physical
spaces.

Bluetooth sensing is based on a discovery process through
which a device can inquire about the presence of other nearby devices.
If those devices are in \emph{discoverable} mode, they will respond
with their Bluetooth addresses (48 bit MAC), and possibly additional
information, such as the device name, the device type (e.g. cellphone
or computer) and available services.

A Bluetooth scanner is a device that periodically scans nearby
Bluetooth devices. Each time a device is detected, it registers and
timestamps that event. Later, this information can be used by external
services or applications. Multiple Bluetooth scanners spread all over
a physical space could thus serve as collection points for Bluetooth
sightings, providing a major tool for observing, recording, modeling
and analyzing that space, physically, digitally and socially
\cite{Oneill:2006vq}.

An infrastructure of this nature can be built from scratch given
the increasingly smaller cost of the technological equipment required.
Alternatively, it can be built upon the large number of Bluetooth
scanners already present in some urban environments. These scanners
are owned by many entities and they serve very diverse purposes, such
as proximity marketing, device localization or OBEX-based interaction.
They could be used, without any additional cost, as nodes in a large
scale collaborative sensing infrastructure. Each node would still scan
for its own purposes, but it would also share part of the generated
data with a central service that would then be able to produce
aggregate information of mutual interest. Both strategies make the
creation of this type of infrastructure relatively simple and feasible
in the short term.

The challenge, however, is how to enable this type of large scale
sensing without creating an overwhelming privacy threat to the users.

% There are already several techniques whose purpose is to ensure
% location privacy in the exchange of information between users and
% service providers. This dissertation aims to find if stochastic
% summarizing techniques are suitable as building blocks for privacy
% preserving systems. To assess their viability, we used them in two
% different Bluetooth scanning scenarios.

\section{Structure and Contributions}
\label{sec:structure}

\subsubsection{Structure}
\label{sec:structure} This dissertation is organized as follows:
Chapter \ref{chap:literature_review} presents some of the related work
in the field of location privacy followed by a brief description of
some algorithms/data structures used in our privacy preserving
approaches. Chapters \ref{cha:gate-counting} and
\ref{cha:causality-tracking} describe both our Bluetooth tracking
scenarios, Collaborative Gate Counting and Causality Tracking, their
requirements and the effectiveness of several stochastic summarizing
techniques in dealing with them. Chapter \ref{cha:conclusion} draws
some general conclusions about our work and presents some future
research directions to be followed.

\subsubsection{Contributions}
\label{sec:contributions}
As a result of the work developed in the context of this dissertation,
two full papers were written. The first one
\cite{gonccalves2011privacy} was published in the OTM 2011 Workshop Proceedings.
The second paper \cite{inforum2012} was published in INFORUM 2012
where it was awarded the prize for best paper in the Mobile and
Ubiquitous Computing session.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../thesis"
%%% End:
